---
title: "Lab Week 8"
author: "Lauren Krohmer"
date: "3/8/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Load Packages

```{r}
library(tidyverse)
library(sf)
library(tmap)
library(leaflet)
library(spatstat)
library(maptools)

```


###Column graph of Texas Oil Spills

```{r}
oil_spills<- read_csv("oil_spills.csv")

df<- oil_spills %>% 
  filter(`Accident State` == "TX" & `Accident Year` <2017) %>% 
  group_by(`Accident Year`) %>% 
  summarise(Loss = sum(`Net Loss (Barrels)`))

colnames(df)<- c("year", "loss")

ggplot(df, aes(x = year, y = loss))+
  geom_col()

```

###Leaflet plot of spill locations in TX in 2016
```{r}

df_loc <- oil_spills %>% 
  filter(`Accident State` == "TX" & `Accident Year` == 2016) %>% 
  select(Latitude, Longitude, `Net Loss (Barrels)`)
colnames(df_loc)<- c("latitude", "longitude", "net_loss")

oil_sf<- st_as_sf(df_loc, coords = c("longitude", "latitude"), crs = 4326)

leaflet(oil_sf) %>% 
  addTiles() %>% 
  addMarkers()

```


## tmap plot with the Texas state shapefile

```{r}

states <- st_read(dsn = ".", layer = "states")

tx_border<- states %>% 
  filter(STATE_NAME == "Texas") %>% 
  st_transform(4326) #change CRS to match other one, would need different code if no previous CRS. in this case we are just changing CRS, not putting new

plot(tx_border)

tm_shape(tx_border)+
  tm_polygons()+
  tm_shape(oil_sf)+ #oil spill points
  tm_dots(size = 0.3)

```

Does this point pattern follow CSR? (even distribution?)

###Convert the data to spatial points patterns (combo of point data and the bounding window)

```{r}
spill_sp<- as(oil_sf, "Spatial") #change to spatial data
spill_ppp <- as(spill_sp, "ppp")
#ppp = point pattern analysis; convert spatial data to point pattern

#bounding window = outline of texas "owin" = window

tx_sp<- as(tx_border, "Spatial")
tx_owin<- as(tx_sp, "owin")

all_ppp <- ppp(spill_ppp$x, spill_ppp$y, window = tx_owin)
#omits points outside of bounding box

#taken spatial info, making in PPP which R recognizes

```

###A density plot:

```{r}

plot(density(all_ppp, sigma =0.5))

```


###Quadrat test for spatial evenness

Are oil spills evenly distributed throughout the state?

```{r}

oil_qt<- quadrat.test(all_ppp, nx = 5, ny = 5)
#5 regions horizontally, 5 vertically - creates quadrats of designated size
oil_qt
#very small p value
#look at data - spatial spread of oil spills look somewhat clustered, not CSR
#quadrat.test NULL is that the data point pattern follows CSR
#small p value - reject NULL - retain alternative that data are NOT CSR - data are NOT evenly distributed


plot(all_ppp)
plot(oil_qt, add = TRUE, cex = 0.4)

#crosses where "illegal" points
#second plot adds quadrat we chose
```


within each quadrat:
-if data were truly spread evenly, the number of events that would occur in each quadrat (above, it is 14.5 oil spill events)
- in not full quadrats, calculates how many events should be given a smaller area (ex. from above is 11.6 total events in upper mid-left quadrat)

If I have data more clustered that CSR - will nearest neighbors be closer or further than in CSR?
- Nearest neighbors will be closer together on average

If data is more even than CSR
- fewer point pairs that have nearest neighbors at short distances from each other

--> consider how all points are surrounding each other with increasing radii around each point to see what the density of enighbors at larger bubbles

###G-function for nearest neighbor analysis

```{r}

r<- seq(0,1, by = 0.01)

oil_gfun<- envelope(all_ppp, fun = Gest, r = r, nsim = 100)

ggplot(oil_gfun, aes(x = r, y = obs))+
  geom_line(color = "black")+
  geom_line(aes(x = r, y = theo), color = "red")
  

#Gest = g estimate
#distance of range to look for nearest neighbors
#nsim= run number of simulations
  #obs is actual data values

#our graph relative to CSR(red) has a higher proporation of point pairs with nearest neighbors at shorter distances compared to CSR data
# on average our data points have closer nearest neighbors than we would expect if completely CSR

```

###Nearest neighbor using Lfunction (Ripley's K, standardized)

```{r}

r2<- seq(0,3, by = 0.5)
oil_lfun <- envelope(all_ppp, fun = Lest, r = r2, nsim = 20, global = TRUE)
#looks at every point in space and makes increasing bubbles until all observation points are incorporated

ggplot(oil_lfun, aes(x = r2, y = obs))+
  geom_line(color = "purple")+
  geom_line(aes(x = r2, y =theo), color = "blue")

```










